---
title: "p8105 homework 2"
author: "Xintao Ding"
date: "9/24/2019"
output: github_document
---

## Problem 1
Read and clean dataset for Mr. Trash Wheel, precipitation 2017 and 2018

```{r import and clean data, warning = FALSE}
library(tidyverse)
library(readxl)

# Mr. Trash Wheel Data
trash_data = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))
trash_data

# 2017 Precipitation Data 
precip2017_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = "2017")

precip2017_data

# 2018 Precipitaion Data
precip2018_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = "2018")

precip2018_data

# Join 2017 and 2018 Precipitation Data
precip_data = bind_rows(precip2017_data, precip2018_data) %>%  
    mutate(month = month.name[month]) 
precip_data = precip_data[c(3,1,2)]
precip_data
```

trash_data is a dataset on dumpster that are sent to a energy facility once it is full. The are `r nrow(trash_data)` obeservations in the trash dataset.Key variables pertains to detailed information on each dumpster collected, these include `r colnames(trash_data)[5:10]`. The median number of sports balls in a dumpster in 2017 was `r median(filter(trash_data, year == 2017)$sports_balls)`.

There are `r nrow(precip_data)` observations in the precipitation dataset. Key variables include `r colnames(precip_data)`. The total precipitation in 2018 was `r sum(precip2018_data$total)`inches. The total precipitation in 2017 was `r sum(precip2017_data$total)`. 

## Problem 2
Read and clean precipiation data for 2017 and 2018
```{r read and clean data }
# import and clean the pols-months data
polsmonth_data = read_csv("./data/fivethirtyeight_datasets/pols-month.csv" ) %>% 
  janitor::clean_names() %>% 
  separate(mon, 
           into = c("year", "month", "day"), 
           convert = TRUE) %>% 
  mutate(
    month = month.abb[month], 
    president = prez_dem - prez_gop, 
    president = recode(president, "-1" = "gop", "1" = "dem")
    ) %>% 
  select(-prez_dem, -prez_gop, -day)

polsmonth_data

# import and clean the snp data
snp_data = read_csv("./data/fivethirtyeight_datasets/snp.csv" ) %>%
  janitor::clean_names() %>% 
  separate(date, 
           into = c("month", "day", "year"), 
           convert = TRUE) %>% 
  mutate(month = month.abb[month]) %>% 
  arrange(year, month) %>% 
  select(year, month, day, close)
snp_data

# import and clean the unemployment data 

unemp_data = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",  
    values_to = "umeployment") %>% 
janitor::clean_names()  
unemp_data

# join three datasets together
snp_pols = full_join(polsmonth_data, snp_data, by= c("year", "month")) 
snp_pols

snps_pols_unemp = inner_join(snp_pols,unemp_data, by = c("year", "month")) 
snps_pols_unemp
```
The pols-month datasets contains `r nrow(polsmonth_data)` observations, and `r ncol(polsmonth_data)` variables. The variables are `r colnames(polsmonth_data)` that describe the number of national politicians who are democratic or repulican during the span of `r range(polsmonth_data$year)`. The variable "president" was created to indicate whether the president is democratic or republican. 

The snp dataset contains `r nrow(snp_data)` observations`, and `r ncol(snp_data)` variables that are related to S&P stock market index during the span of `r range(snp_data$year)`. The variable "close" is the  closing value of S&P index on the associated date. 

The unemployment data contains`r nrow(unemp_data)` observations, and `r ncol(unemp_data)` variables. The variables are `r colnames(unemp_data)` that documents percentage of unemployment in each of the 12 months of the associated year during the span of `r range(unemp_data$year)` 

## Problem 3
Import, clean, and plot popular baby name data
```{r popular baby names}
# import and clean baby name data
name_data = read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity),
    childs_first_name = str_to_title(childs_first_name)
  ) %>% 
  distinct()

name_data 

# create Olivia popularity table

Olivia = filter(name_data, childs_first_name == "Olivia" ) %>% 
  select(ethnicity, year_of_birth, rank) %>% 
  arrange(ethnicity)
Olivia

# create male popular name table

male_name = filter(name_data, gender == "male" ) %>% 
  select(childs_first_name, ethnicity, year_of_birth, rank) %>% 
  arrange(childs_first_name)
male_name

# plotting white non-hispanic male born in 2016' names vs. rank
white_male_2016 = filter(name_data, gender == "male", ethnicity == "white non hispanic", year_of_birth == "2016") %>% 
  ggplot(aes(x = rank , y = count)) + geom_point() + labs(title = "White Non-Hispanic Male Born in 2016 Number of Names vs, Rank", x = "rank in popularity", y = "number of children with a name")

white_male_2016
```